Inequality in consumption

### 2017-03-08 inequality-resistance/inequality-resistance.Rmd
many changes

```{r load_groups}
resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
consumption_groups = read_tsv('../states/resistance/consumption_groups.tsv')
```

```{r load_antibiogram}
organisms = c('Escherichia coli')

@@

consumption = usage %>%
  left_join(consumption_groups, by='drug') %>%
  group_by(bene_id, drug_group) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup

@@

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}
```

>>>>

```{r load_groups}
resistance_groups = read_tsv('../states/resistance/resistance_groups.tsv')
consumption_groups = read_tsv('../states/resistance/consumption_groups.tsv')
drug_groups = resistance_groups$drug_group
```

```{r load_antibiogram}
organisms = c('Escherichia coli',
              'Proteus mirabilis', 'Enterobacter cloacae',
              'Enterococcus faecium',
              'Coagulase-negative staphylococci',
              'Methicillin-resistant S. aureus (MRSA)',
              'Methicillin-susceptible S. aureus (MSSA)',
              'Streptococcus pneumoniae (non-meningitis)',
              'Staphylococcus aureus')

@@

consumption = usage %>%
  left_join(consumption_groups, by='drug') %>%
  group_by(bene_id, drug_group) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ungroup %>%
  complete(bene_id, drug_group, fill=list('n_claims'=0L)) %>%
  left_join(denom, by='bene_id')

@@

normalize = function(x) x / max(x)
juran = function(x) {
  # better to do with percent_rank and cume_dist
  fraction_consumed = cumsum(sort(x, decreasing=TRUE)) %>% normalize
  fraction_users = 1:length(x) %>% normalize
  fraction_nonzero(fraction_consumed + fraction_users - 1)
}

klasma = function(x, threshold) {
  if (!between(threshold, 0.0, 1.0)) stop("bad threshold")
  data_frame(x=sort(x, decreasing=TRUE)) %>%
    mutate(csx=cumsum(x), cd=cume_dist(csx)) %$%
    sum(x[cd <= 0.1]) / sum(x)
}
```

```{r}
mean_abg = abg %>%
  group_by(bug, drug_group, state) %>%
  summarize(mpr=mean(percent_nonsusceptible),
            wmpr=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)))

state_consumption = consumption %>%
  group_by(state, drug_group) %>%
  summarize(f_nz=fraction_nonzero(n_claims),
            pct_nz=100*f_nz,
            juran0=juran(n_claims),
            juran_nz=juran(nonzero(n_claims)),
            gini0=gini(n_claims),
            gini_nz=gini(nonzero(n_claims)),
            k10_nz=klasma(nonzero(n_claims), 0.1),
            n_bene=n(),
            n_claims=sum(n_claims),
            cpkp=n_claims/n_bene*1000,
            cpku=cpkp/f_nz)

```

```{r do_models}
models = function(x, independent) {
  univariate_formula = paste(independent, '~ cpkp') %>% formula
  univariate_results = lm(univariate_formula, data=x) %>%
    anova %>% tidy %>% mutate(model='univariate')
  multivariate_formula = paste(independent, '~ pct_nz + cpku') %>% formula
  multivariate_results = lm(multivariate_formula, data=x) %>%
    anova %>% tidy %>% mutate(model='multivariate')
  bind_rows(univariate_results, multivariate_results)
}

results = left_join(mean_abg, state_consumption, by=c('state', 'drug_group')) %>%
  group_by(bug, drug_group) %>%
  do(models(., 'mpr')) %>%
  select(bug, drug_group, model, term, estimate, p.value)

results2 = left_join(abg, state_consumption, by=c('state', 'drug_group')) %>%
  group_by(bug, drug_group) %>%
  do(models(., 'percent_nonsusceptible')) %>%
  select(bug, drug_group, model, term, estimate, p.value)
```

# Results

## Correlations between metrics

In general,

- CPKP correlates strongly with `f_nz` (makes sense, since most users take ~1 rx)
- Intensive metrics correlate with one another (`gini_nz` and `juran_nz`)
- Extensive metrics correlate (e.g., `juran` with `f_nz`)
- Intensive and extensive metrics correlation (`juran_nz` with `f_nz`)

Thus, in general, increased consumption goes with increased inequality.

## Klasma 10

Klasma 10 is the fraction of the drugs consumed by the top 10% of users. Consider bug/drug combos where `k10_nz` in the model `percent_susceptible` (or mean r) `~ pct_nz + k10_nz` is significant. Those are all either tmp-smx or tetracycline, and in each case, the estimate for `cpkp` in the univariate model matches the sign of `pct_nz` in the multivariate model, and `k10_nz` in the multivariate model is always negative.

## Using CPKU

I tried to use CPKU as an intensive measure: what's the average consump (CPK) among users (U)? I got similar results to when I was using other intensive metrics.

## f_nz vs. CPKP

In general, the fraction nonzero is a better predictor of resistance than CPKP, i.e., the extensive metric is better than the garbled intensive metric.

### 2017-03-08 inequality/consumption-inequality.Rmd
many changes

>>>>

```{r inequality_by_region2}
# This uses code from inequality-resistance
normalize = function(x) x / max(x)
cumrelfreq = function(df, var) {
  n = sort(df[[var]], decreasing=FALSE)
  data_frame(x=cume_dist(n), y=normalize(cumsum(n))) %>%
    group_by(x) %>%
    summarize(y=max(y))
}

m = consumption_of('macrolide')

x = m %>%
  left_join(census_regions, by='state') %>%
  group_by(region) %>%
  do(cumrelfreq(., 'n_claims'))

p = x %>%
  ggplot(aes(x=x, y=y, color=region, group=region)) +
  geom_line() + geom_point() +
  xlab('cumulative fraction of users') +
  ylab('cumulative fraction of drugs consumed') +
  theme_minimal()

show(p)
```
